\documentclass[11pt]{article}

\input{macros}

\usepackage[margin=1in]{geometry}

\title{Vegas: A Typed Game Calculus\\with Commit-Reveal Information Structure}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We present \emph{Vegas}, a typed domain-specific calculus for expressing
strategic games as programs. Vegas programs compose five constructs---return,
let-binding, sampling, commit, and reveal---under a visibility-aware static
discipline. Each variable binding carries a visibility annotation---public or
hidden with a designated owner---and a view-scoping judgment ensures
that commit guards depend only on the committing player's view. We give Vegas four semantic representations---denotational,
trace, action graph, and native---connected by adequacy, normalization,
execution uniqueness, and commutativity theorems. A strategic bridge packages
any Vegas program as a kernel game whose strategy spaces are view-based
behavioral strategies, with a proof that expected utility is
preserved---enabling game-theoretic solution concepts (Nash equilibria,
dominance) to be applied directly to programs. We further compile Vegas programs to
Multi-Agent Influence Diagrams (MAIDs) and prove that compilation preserves
outcome distributions. The compiled MAID has perfect recall, enabling a
Vegas Kuhn theorem: every product mixture over pure strategies is
outcome-equivalent to some behavioral strategy profile. The entire
development is mechanized in Lean~4, comprising approximately 10,000 lines
of verified code for the Vegas calculus and 33,000 lines for a supporting
game theory library.
\end{abstract}

%% =====================================================================
\section{Introduction}\label{sec:intro}
%% =====================================================================

Strategic interaction pervades computing: auctions, mechanism design,
multi-agent systems, and smart contracts all require reasoning about programs
whose outcomes depend on the choices of multiple self-interested agents. Yet
the tools available to formalize such systems remain fragmented. Game theory
provides solution concepts (Nash equilibria, subgame perfection) but lacks the
compositional, syntactic structure of programming languages. Conversely,
programming language theory offers rich type systems and verified semantics but
rarely accounts for the strategic dimension of computation.

\paragraph{The information problem.}
The central challenge in modeling strategic interaction is \emph{imperfect
information}: players make decisions based on partial observations of the
game state. In extensive-form games, this is modeled by information sets. In
multi-agent influence diagrams (MAIDs), it is captured by parent edges in a
directed acyclic graph. Both representations encode information structure as a
\emph{side condition} external to the game's operational content. This
separation makes it easy to write games with inconsistent information
structure---a player's strategy accidentally depending on variables they should
not observe---with the error caught only by manual inspection.

\paragraph{Vegas: information structure through types.}
We propose \emph{Vegas}, a calculus with visibility-annotated contexts and a
type-state-like scoping discipline for information structure. Every variable binding carries a
\emph{visibility annotation}: either $\pub\;\tau$ (publicly observable) or
$\hid(i, \tau)$ (hidden, owned by player~$i$). The static discipline ensures:
%
\begin{itemize}[nosep]
\item Payoff expressions may only reference public variables.
\item A commit guard (the Boolean condition constraining a player's choice)
      may reference the committing player's \emph{view}: all public variables
      plus that player's own hidden variables.
\item Reveal makes a hidden variable public, modeling information disclosure.
\end{itemize}
%
This design draws inspiration from the commit-reveal pattern in blockchain
protocols, where a player first commits to a value (publishing a hash) and
later reveals it. In Vegas, commit and reveal are first-class language
constructs whose information-flow properties are enforced by visibility-aware
scoping rather than by cryptographic mechanisms.

\paragraph{Contributions.}
\begin{enumerate}[nosep]
\item \textbf{The Vegas calculus} (\Cref{sec:overview,sec:calculus}): a typed
      DSL with five constructs---$\Ret$, $\Let$, $\Sample$, $\Commit$,
      $\Reveal$---and a visibility-aware context system that enforces
      information structure through a separate view-scoping judgment.

\item \textbf{Four semantic representations} (\Cref{sec:semantics}):
      a denotational semantics ($\outcomeDist$) producing weighted outcome
      distributions; a trace semantics enumerating execution paths;
      an action graph semantics providing order-independent DAG-based
      execution; and a native intermediate semantics bridging to MAID
      evaluation.
      Proofs connecting these layers include
      \emph{normalization} (distributions sum to~1),
      \emph{adequacy} (denotational = trace),
      \emph{support correctness} (reachable outcomes = support),
      \emph{execution uniqueness} (action graph order-independence),
      and \emph{commutativity} (adjacent independent constructs commute).

\item \textbf{Strategic bridge} (\Cref{sec:strategic}): a construction
      $\toKG(p, \rho)$ packaging any Vegas program as a kernel game whose
      strategies are view-based behavioral strategies---functions from a
      player's visible state to action distributions. The typing discipline
      guarantees that strategies cannot depend on unobserved state, unlike
      in extensive-form games where this is a convention. An EU preservation
      theorem connects the kernel game's expected utility to the Vegas
      denotational semantics.

\item \textbf{MAID compilation and Kuhn theorem} (\Cref{sec:maid}): a
      verified compiler from Vegas programs to Multi-Agent Influence Diagrams,
      with a correctness theorem showing that the compiled MAID's outcome
      distribution equals the Vegas denotational semantics. The compiled MAID
      has perfect recall, which yields a Vegas Kuhn theorem: every product
      mixture over pure strategies is outcome-equivalent to some behavioral
      strategy profile.

\item \textbf{Lean~4 mechanization} (\Cref{sec:mechanization}): the entire
      development is formalized in Lean~4, including ${\sim}10{,}000$ lines for
      the Vegas calculus and ${\sim}33{,}000$ lines for a supporting game
      theory library providing extensive-form games, normal-form games, MAIDs,
      and their interrelations.
\end{enumerate}

\paragraph{Paper outline.}
\Cref{sec:overview} introduces Vegas through examples.
\Cref{sec:calculus} defines the calculus formally.
\Cref{sec:semantics} presents the four semantic representations and their
connecting theorems.
\Cref{sec:strategic} describes the strategic bridge.
\Cref{sec:maid} covers MAID compilation.
\Cref{sec:mechanization} discusses the mechanization.
\Cref{sec:related} surveys related work.
\Cref{sec:conclusion} concludes.


%% =====================================================================
\section{Overview by Example}\label{sec:overview}
%% =====================================================================

We introduce Vegas through three games that illustrate the key features
of the calculus: visibility annotations, commit-reveal, guards, and
information flow.

\subsection{Matching Pennies}

In Matching Pennies, two players simultaneously choose heads or tails. Player~0
wins if the choices match; Player~1 wins if they differ. The ``simultaneous''
aspect means neither player observes the other's choice before committing.

In Vegas, we express this as:
%
\begin{quote}
\begin{tabbing}
\qquad\= \qquad\= \kill
$\Commit\;a\;\mathbf{by}\;\pl{0}\;\mathbf{where}\;\mathit{true}$ \\
$\Commit\;b\;\mathbf{by}\;\pl{1}\;\mathbf{where}\;\mathit{true}$ \\
$\Reveal\;a'\;\mathbf{from}\;a$ \\
$\Reveal\;b'\;\mathbf{from}\;b$ \\
$\Ret\;[\pl{0} \mapsto \mathbf{if}\;a'=b'\;\mathbf{then}\;1\;\mathbf{else}\;{-1},$ \\
\> $\pl{1} \mapsto \mathbf{if}\;a'=b'\;\mathbf{then}\;{-1}\;\mathbf{else}\;1]$
\end{tabbing}
\end{quote}
%
The commit construct $\Commit\;a\;\mathbf{by}\;\pl{0}\;\mathbf{where}\;\mathit{true}$ introduces a hidden variable~$a$ owned by Player~0. Since the committed type is Boolean, the available choices are the two Boolean values; the guard $\mathit{true}$ imposes no additional restriction. After both players commit, the program reveals both values and computes payoffs from the now-public variables $a'$ and $b'$.

\paragraph{Visibility in action.}
After the two commits, the visibility context is:
\[
  \Gamma = [\,(b, \hid(\pl{1}, \bool)),\; (a, \hid(\pl{0}, \bool))\,]
\]
Player~0's view $\Gamma\!\!\downarrow_{\pl{0}} = [(a, \hid(\pl{0}, \bool))]$
contains only their own hidden variable---not Player~1's. This enforces the
simultaneous nature of the game: neither player's strategy can depend on the
other's choice. After both reveals, $a'$ and $b'$ are $\pub$, satisfying
the typing requirement that payoffs reference only public variables.

\subsection{Sequential Reveal}

Not all games are simultaneous. Consider a variant where Player~0 commits
and reveals before Player~1 moves---a simple model of sequential play:
%
\begin{quote}
\begin{tabbing}
\qquad\= \qquad\= \kill
$\Commit\;a\;\mathbf{by}\;\pl{0}\;\mathbf{where}\;\mathit{true}$ \\
$\Reveal\;a'\;\mathbf{from}\;a$ \\
$\Commit\;b\;\mathbf{by}\;\pl{1}\;\mathbf{where}\;\mathit{true}$ \\
$\Reveal\;b'\;\mathbf{from}\;b$ \\
$\Ret\;[\pl{0} \mapsto \mathbf{if}\;a'=b'\;\mathbf{then}\;1\;\mathbf{else}\;{-1},$ \\
\> $\pl{1} \mapsto \mathbf{if}\;a'=b'\;\mathbf{then}\;{-1}\;\mathbf{else}\;1]$
\end{tabbing}
\end{quote}
%
The structure is the same as Matching Pennies except that $a$ is revealed
\emph{before} Player~1 commits. Both guards are trivially~$\mathit{true}$,
so the program satisfies all standing assumptions (\Cref{def:standing}).

\paragraph{Sequential information flow.}
At the point of Player~1's commit, the context is
$[(a', \pub\;\bool),\; (a, \hid(\pl{0}, \bool))]$. Player~1's view includes
$a'$ (public), so Player~1's commit kernel may depend on the revealed value.
This is the key difference from Matching Pennies, where both commits precede
all reveals and neither player's view includes the other's variable.

The commit-reveal pattern naturally models both simultaneous (commit before any
reveal) and sequential (interleave commits and reveals) information structures
within the same calculus. The interleaving order---not the textual order of
constructs---determines the information available to each player. Our
commutativity theorems (\Cref{sec:semantics}) confirm that reordering
independent constructs does not affect semantics.

\subsection{Conditioned Choices}

Commit guards can restrict a player's available actions based on observable
state. Consider a variant where Player~0's revealed choice becomes the guard
for Player~1:
%
\begin{quote}
\begin{tabbing}
\qquad\= \qquad\= \kill
$\Commit\;a\;\mathbf{by}\;\pl{0}\;\mathbf{where}\;\mathit{true}$ \\
$\Reveal\;a'\;\mathbf{from}\;a$ \\
$\Commit\;b\;\mathbf{by}\;\pl{1}\;\mathbf{where}\;a'$ \\
$\Reveal\;b'\;\mathbf{from}\;b$ \\
$\Ret\;[\pl{0} \mapsto 1,\; \pl{1} \mapsto 0]$
\end{tabbing}
\end{quote}
%
Player~1's guard is the variable $a'$, which is Player~0's revealed choice
interpreted as a Boolean. When $a' = \mathit{true}$, Player~1's choices are
unrestricted (both Boolean values satisfy the guard). But when
$a' = \mathit{false}$, \emph{no} Boolean value satisfies the guard---a
deadlock. This program violates the $\Legal$ condition
(\Cref{def:standing}), which requires that for every erased environment, at
least one action satisfies the guard. The example illustrates both the
expressiveness of non-trivial guards and the necessity of the legality
assumption.

\paragraph{View scoping of guards.}
The guard $a'$ is well-scoped: it depends only on $a'$, which is public
and hence in Player~1's view at the point of the commit. A guard that
referenced a variable hidden from the committing player would violate the
view-scoping judgment (\Cref{def:viewscoped}).

\subsection{Nature and Sampling}

\paragraph{Missing:} \textit{The calculus includes a $\Sample$ construct with
three sampling modes---public nature, private nature, and player-private---that
draw from distribution expressions and bind the result with appropriate
visibility. A small example (e.g., Bayesian game with private types drawn from
a prior) should be added here to complete the tour of language features.}


%% =====================================================================
\section{The Vegas Calculus}\label{sec:calculus}
%% =====================================================================

The Vegas calculus is parameterized over a set of players $\Players$ and an
expression language interface $L$, and operates on visibility-aware contexts
that track which variables each player can observe.

\subsection{Expression Language Interface}

The calculus is generic over an expression language interface $L$
(\leanref{IExpr}) that provides types, values, expression syntax with an
evaluator, distribution expressions, and a dependency-tracking discipline.
All Vegas metatheory is proved once for any $L$; a concrete instantiation with
integers, Booleans, arithmetic, equality, and conditionals is used for
examples. The full interface is described in \Cref{sec:mechanization}.

\subsection{Visibility Contexts}\label{sec:vctx}

\begin{definition}[Visibility-Annotated Types]
A \emph{binding type} $\beta$ is either public or hidden:
\[
  \beta \;\;::=\;\; \pub\;\tau \;\mid\; \hid(i, \tau)
\]
where $\tau : L.\Ty$ and $i \in \Players$. We write $|\beta|$ for the
\emph{base type}: $|\pub\;\tau| = |\hid(i,\tau)| = \tau$.
\end{definition}

\begin{definition}[Visibility Context]
A \emph{visibility context} $\Gamma$ is an ordered list of bindings
$(x, \beta)$ where $x$ is a variable identifier and $\beta$ is a binding type:
\[
  \Gamma \;\;::=\;\; [] \;\mid\; (x, \beta) :: \Gamma
\]
\end{definition}

\paragraph{Context projections.}
Given $\Gamma$, we define projections (\Cref{fig:projections}):

\begin{figure}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lp{7.5cm}@{}}
\toprule
Notation & Definition \\
\midrule
$\Gamma\!\!\downarrow_i$ &
  Filter $\Gamma$ to bindings visible to player~$i$: keep $(x, \pub\;\tau)$
  and $(x, \hid(i, \tau))$, drop $(x, \hid(j, \tau))$ for $j \ne i$. \\
$\Gamma^{\pub}$ &
  Keep only $\pub$ bindings. \\
$\lfloor\Gamma\rfloor$ &
  Replace every $\hid(i,\tau)$ with $\pub\;\tau$. \\
$\|\Gamma\|$ &
  Erase visibility annotations: $(x, \beta) \mapsto (x, |\beta|)$.
  Produces a plain context over $L.\Ty$. \\
$\|\Gamma^{\pub}\|$ &
  Erase, keeping only public variables. \\
\bottomrule
\end{tabular}
\caption{Context projections. Each maps a visibility context to a derived
context consumed by a different part of the calculus.}\label{fig:projections}
\end{figure}

\begin{example}[Context Projections]\label{ex:projections}
After the two commits in Matching Pennies, the context is
$\Gamma = [(b, \hid(\pl{1}, \bool)),\; (a, \hid(\pl{0}, \bool))]$.
Then:
\begin{align*}
  \Gamma\!\!\downarrow_{\pl{0}}
    &= [(a, \hid(\pl{0}, \bool))] \\
  \Gamma^{\pub}
    &= [] \\
  \|\Gamma\|
    &= [(b, \bool),\; (a, \bool)] \\
  \|\Gamma^{\pub}\|
    &= []
\end{align*}
After both reveals extend $\Gamma$ with $(b', \pub\;\bool)$ and
$(a', \pub\;\bool)$, the erasure $\|\Gamma'^{\pub}\| = [(b', \bool), (a', \bool)]$
provides the context for payoff expressions.
\end{example}


\subsection{Syntax}\label{sec:syntax}

A Vegas program $\Gamma \vdash p$ in context $\Gamma$ is given by the
grammar in \Cref{fig:grammar}.

\begin{figure}[ht]
\centering
\fbox{\parbox{0.92\textwidth}{
\begin{align*}
p \;\;::=\;\;
  & \Ret\;\mathit{payoffs}
    & \text{where } \mathit{payoffs} : \textstyle\prod_{i \in \Players}
      L.\Expr(\|\Gamma^{\pub}\|, \intt) \\
  \mid\;\;
  & \Let\;x : \tau = e \;\mathbf{in}\; k
    & \text{where } e : L.\Expr(\|\Gamma^{\pub}\|, \tau),\;
      k : (x{:}\pub\;\tau) :: \Gamma \vdash p \\
  \mid\;\;
  & \Sample\;x : \beta \sim_m D \;\mathbf{in}\; k
    & \text{where } D : L.\mathit{DistExpr}(\|\mathit{ctx}_m(\Gamma)\|, |\beta|),\;
      k : (x{:}\beta) :: \Gamma \vdash p \\
  \mid\;\;
  & \Commit\;x\;\mathbf{by}\;i\;\mathbf{where}\;R\;\mathbf{in}\;k
    & \text{where } R : L.\Expr((x{:}\tau) :: \|\Gamma\|,\;\bool) \\
  & & k : (x{:}\hid(i,\tau)) :: \Gamma \vdash p \\
  \mid\;\;
  & \Reveal\;y\;\mathbf{from}\;x\;\mathbf{in}\;k
    & \text{where } (x{:}\hid(i,\tau)) \in \Gamma,\;
      k : (y{:}\pub\;\tau) :: \Gamma \vdash p
\end{align*}
}}
\caption{Vegas syntax. The sampling context $\mathit{ctx}_m(\Gamma)$ depends on
the mode $m$: $\Gamma^{\pub}$ for public/private nature,
$\lfloor \Gamma\!\!\downarrow_i \rfloor$ for player-private.}\label{fig:grammar}
\end{figure}

\noindent
The five constructs operate as follows:
\begin{itemize}[nosep]
\item $\Ret$: Terminate with per-player payoffs, each typed over
      $\|\Gamma^{\pub}\|$.
\item $\Let$: Bind a deterministic expression to a fresh $\pub$ variable.
\item $\Sample$: Draw from a distribution and bind the result.
      The sampling \emph{mode} $m$ determines both the visibility of $x$ and the
      context available to $D$: public nature ($x$ is $\pub$, $D$ sees
      $\Gamma^{\pub}$), private nature ($x$ is $\hid$, $D$ sees
      $\Gamma^{\pub}$), or player-private ($x$ is $\hid(i,\tau)$, $D$ sees
      $\lfloor\Gamma\!\!\downarrow_i\rfloor$).
\item $\Commit$: Player~$i$ chooses a value of type~$\tau$ subject
      to guard~$R$. In the raw core, the guard is typed over the full erased
      context; the separate view-scoping judgment requires its dependencies to
      lie in the proposed value and the player's view. The result is bound as
      $\hid(i, \tau)$.
\item $\Reveal$: Make a hidden variable public by binding a fresh $\pub$ copy.
      Reveal is an explicit construct so that the \emph{timing} of information
      disclosure is under program control---a commit followed by an immediate
      reveal models open play, while deferring the reveal models sealed
      commitment.
\end{itemize}


\subsection{Well-Formedness}\label{sec:wf}

\begin{definition}[Well-Formedness]\label{def:wf}
A program $\Gamma \vdash p$ is \emph{well-formed}, written $\WF(p)$, if three
conditions hold:
\begin{itemize}[nosep]
\item \textbf{Fresh bindings}: every variable introduced by $\Let$, $\Sample$,
      $\Commit$, or $\Reveal$ is fresh in the current context, enforcing SSA.
\item \textbf{Reveal-completeness}: every committed secret is eventually
      revealed exactly once.
\item \textbf{View scoping}: every commit guard depends only on the proposed
      value and variables visible to the committing player.
\end{itemize}
\end{definition}

\begin{definition}[Reveal-Completeness]\label{def:revealcomplete}
$\RevealComplete(\mathit{pending}, p)$ is defined by structural recursion:
\begin{itemize}[nosep]
\item $\RevealComplete(\mathit{pending}, \Ret\;\cdot) \iff
      \mathit{pending} = []$
\item $\RevealComplete(\mathit{pending}, \Let\;x = e\;\mathbf{in}\;k) \iff
      \RevealComplete(\mathit{pending}, k)$
\item $\RevealComplete(\mathit{pending}, \Sample\;x \sim D\;\mathbf{in}\;k) \iff
      \RevealComplete(\mathit{pending}, k)$
\item $\RevealComplete(\mathit{pending}, \Commit\;x\;\cdot\;\mathbf{in}\;k) \iff
      \RevealComplete(x :: \mathit{pending}, k)$
\item $\RevealComplete(\mathit{pending}, \Reveal\;y\;\mathbf{from}\;x\;\mathbf{in}\;k) \iff
      x \in \mathit{pending} \;\wedge\;
      \RevealComplete(\mathit{pending} \setminus \{x\}, k)$
\end{itemize}
A program is \emph{reveal-complete} if $\RevealComplete([], p)$ holds: every
committed secret is revealed exactly once before termination.
\end{definition}

\begin{definition}[View Scoping]\label{def:viewscoped}
For a commit site $\Commit\;x\;\mathbf{by}\;i\;\mathbf{where}\;R$, define
\[
  \mathsf{ScopedGuard}(i, x, \Gamma, R)
  \;\;\iff\;\;
  \mathsf{deps}(R) \subseteq \{x\} \cup \mathsf{Vis}_i(\Gamma).
\]
The recursive judgment $\mathsf{Scoped}(p)$ requires this condition at every
commit site. This is the formal statement that guards may mention only the
committer's view, even though the raw syntax types them over the full erased
context. In the mechanization, $\mathsf{deps}$ is provided by the expression
interface and justified by dependency-soundness theorems.
\end{definition}


\subsection{Standing Assumptions}\label{sec:assumptions}

\begin{definition}[Standing Assumptions]\label{def:standing}
Let $p$ be a Vegas program and $\sigma$ an operational profile. The following
predicates are defined by structural recursion on~$p$:
\begin{itemize}[nosep]
\item $\Legal(p)$: At every commit site, for every erased environment
      $\rho : \Env(\|\Gamma\|)$, there exists some value
      $a : L.\Val(\tau)$ with $R[x \mapsto a, \rho] = \mathit{true}$.
      This is a feasibility condition: the game never deadlocks.
      (See the conditioned-choice example in \Cref{sec:overview} for a
      program that violates it.)

\item $\NormDists(p)$: At every sample site, for every environment $\rho$,
      $\sum_{v} D(\rho)(v) = 1$. Used by the normalization theorem
      (\Cref{thm:norm}).

\item $\NormOn(\sigma, p)$: At every commit site, for every erased environment
      $\rho$, $\sum_{v} \sigma(i, x, R)(\rho)(v) = 1$. Used together with
      $\NormDists$ for normalization.
\end{itemize}
\end{definition}

\noindent
Together with $\WF(p)$ and $\RevealComplete([], p)$, these form the
preconditions for our main theorems.

\begin{definition}[Fair Play]\label{def:admissible}
An operational profile~$\sigma$ is \emph{fair play} for~$p$ if at every commit
site $(x, i, R)$ and every erased environment $\rho$, the kernel
$\sigma(i, x, R)(\rho)$ is supported on values
$a$ satisfying $R[x \mapsto a, \rho] = \mathit{true}$---i.e., players only
choose actions that satisfy the guard. Fair play connects strategy properties
to trace properties: under a fair-play profile, every trace with positive
weight is legal (\Cref{thm:fairplay}).
\end{definition}


\subsection{Strategies}\label{sec:strategies}

\begin{definition}[Operational Commit Kernel]
An \emph{operational commit kernel} at a commit site with owner~$i$ in context
$\Gamma$ is a function
\[
  \kappa : \Env(\|\Gamma\|) \to \FDist{L.\Val(\tau)}
\]
mapping the current erased environment to a distribution over values of the
committed type~$\tau$. This is the evaluator-facing notion used by the
denotational and trace semantics.
\end{definition}

\begin{definition}[Operational Profile]
An \emph{operational profile} $\sigma$ provides an operational commit kernel for
every commit site: given a context $\Gamma$, player $i$, variable $x$, and
guard $R$, it returns
\[
  \sigma(i, x, R) : \Env(\|\Gamma\|) \to \FDist{L.\Val(\tau)}.
\]
\end{definition}

\begin{definition}[View-Based Strategies]
Let $\mathcal{V}_i(\Gamma) = \Env(\|\Gamma\!\!\downarrow_i\|)$.
\begin{itemize}[nosep]
\item A \emph{pure strategy} for player~$i$ maps each commit site in context
      $\Gamma$ to a deterministic choice function
      $\mathcal{V}_i(\Gamma) \to L.\Val(\tau)$.
\item A \emph{behavioral strategy} for player~$i$ maps each commit site to a
      normalized kernel
      $\mathcal{V}_i(\Gamma) \to \FDist{L.\Val(\tau)}$.
\item A \emph{mixed strategy} is a probability distribution over pure
      strategies.
\end{itemize}
Joint pure, behavioral, and mixed profiles are defined pointwise over players.
Behavioral profiles are converted to operational profiles by projecting each
erased environment to the corresponding player view
(\leanref{projectViewEnv}). This is the bridge from the game-theoretic
strategy layer to the operational semantics.
\end{definition}

\paragraph{Information discipline by construction.}
In extensive-form games, a behavioral strategy is a function from
\emph{information sets} to action distributions, but nothing in the type
prevents defining a strategy that conditions on the full game history---the
information-set constraint is enforced by convention.
In Vegas, the strategy type $\mathcal{V}_i(\Gamma) \to \FDist{L.\Val(\tau)}$
makes the constraint structural: a player's strategy is a function from
\emph{only the variables the player can observe}, as determined by the
visibility context. The view projection $\Gamma\!\!\downarrow_i$ erases
variables hidden from player~$i$, so there is no way to write a strategy that
accidentally depends on unobserved state. This is the central payoff of the
visibility-aware type system: information structure is enforced by
construction, not by side conditions. In the mechanization, we prove that
operationalized behavioral strategies automatically satisfy the
observation-respecting property (\leanref{toOperationalProfile\_respectsObservation}).


%% =====================================================================
\section{Semantics}\label{sec:semantics}
%% =====================================================================

Vegas has two primary semantic representations---denotational and
trace---connected by adequacy. An action graph formalization establishes
order-independence, and a native intermediate semantics bridges to MAID
evaluation (\Cref{sec:maid}). \Cref{fig:layers} summarizes the architecture.

\begin{figure}[ht]
\centering
\fbox{\parbox{0.88\textwidth}{\centering
\begin{tabular}{ccc}
  \textbf{Denotational} & & \textbf{Trace} \\
  $\densub{p}{\sigma}(\rho) : \FDist{\Outcome}$
  & $\xleftarrow{\;\text{adequacy}\;}$ &
  $\sum_{t} \trweight{\sigma}(t, \rho) \cdot \delta_{\trout(t,\rho)}$ \\[8pt]
  $\downarrow$ \text{\small normalization} & & $\downarrow$ \text{\small support = Reach} \\[4pt]
  $\PMF{\Outcome}$ & & \\[4pt]
  $\downarrow$ \text{\small $\toKG$} & & \\[4pt]
  \textbf{Kernel Game} $\KG$ & $\xrightarrow{\;\text{EU pres.}\;}$ &
  Nash / dominance \\[12pt]
  \multicolumn{3}{c}{\textbf{Action Graph}: $\mathit{extractEvents}(p) \to
    \mathsf{Execution}$ \quad (\text{order-independence})}
\end{tabular}
}}
\caption{Semantic architecture. The denotational and trace semantics are
connected by adequacy. Normalization promotes to PMF, enabling the strategic
bridge $\toKG$ (\Cref{sec:strategic}). The action graph
(\Cref{sec:actiongraph}) provides a complementary order-independent execution
model. MAID compilation (\Cref{sec:maid}) proceeds via a separate native
intermediate semantics proved equal to $\outcomeDist$.}\label{fig:layers}
\end{figure}


\subsection{Weighted Distributions}

The type $\FDist{\alpha}$ denotes finitely-supported functions
$\alpha \to \NNQ$ (nonnegative rationals). This is a weighted distribution
monad with:
\begin{itemize}[nosep]
\item $\mathrm{pure}(a)(x) = \begin{cases} 1 & \text{if } x = a \\ 0 & \text{otherwise}\end{cases}$
\item $(d \bind f)(y) = \sum_{x} d(x) \cdot f(x)(y)$
\item $\mathrm{totalWeight}(d) = \sum_{x} d(x)$
\item $\mathrm{support}(d) = \{x \mid d(x) \ne 0\}$
\end{itemize}
When $\mathrm{totalWeight}(d) = 1$, the distribution can be promoted to a
probability mass function $\PMF{\alpha}$.


\subsection{Denotational Semantics}\label{sec:densem}

The denotational semantics is defined by structural recursion on programs
(\Cref{fig:outcomedist}).

\begin{figure}[ht]
\centering
\fbox{\parbox{0.92\textwidth}{
\[
\densub{p}{\sigma}(\rho) \;\coloneqq\;
\begin{cases}
  \mathrm{pure}(\mathit{evalPayoffs}(\mathit{payoffs}, \rho))
    & \text{if } p = \Ret\;\mathit{payoffs} \\[6pt]
  \densub{k}{\sigma}(\rho[x \mapsto L.\mathit{eval}(e, \rho^{\pub})])
    & \text{if } p = \Let\;x = e\;\mathbf{in}\;k \\[6pt]
  L.\mathit{evalDist}(D, \rho_m) \bind
    (\lambda v.\; \densub{k}{\sigma}(\rho[x \mapsto v]))
    & \text{if } p = \Sample\;x \sim_m D\;\mathbf{in}\;k \\[6pt]
  \sigma(i,x,R)(\rho) \bind
    (\lambda v.\; \densub{k}{\sigma}(\rho[x \mapsto v]))
    & \text{if } p = \Commit\;x\;\cdot\;\mathbf{by}\;i\;\cdot\;\mathbf{in}\;k \\[6pt]
  \densub{k}{\sigma}(\rho[y \mapsto \rho(x)])
    & \text{if } p = \Reveal\;y\;\mathbf{from}\;x\;\mathbf{in}\;k
\end{cases}
\]
}}
\caption{Denotational semantics $\densub{p}{\sigma}(\rho) : \FDist{\Outcome}$.
Here $\rho^{\pub}$ projects to public variables, $\rho_m$ projects according
to sampling mode, and $\rho[x \mapsto v]$ extends the environment. For the
game-theoretic strategy layer, operational kernels are obtained from
view-indexed behavioral strategies by first projecting $\rho$ to the player's
visible environment.}\label{fig:outcomedist}
\end{figure}

\noindent
The outcome type $\Outcome = \Players \to_{\mathrm{fin}} \mathbb{Z}$ is a
finitely-supported map from players to integer payoffs.
Weights multiply along paths: if a sample distribution assigns weight~$w$ to a
value and the profile kernel assigns weight~$w'$ to a choice, the combined
weight is $w \cdot w'$.


\subsection{Trace Semantics}\label{sec:tracesem}

\begin{definition}[Trace]\label{def:trace}
A \emph{trace} $t \in \Tr{p}$ is an inductive record indexed by the program it
traverses. At each branching point, the trace records the chosen value:
\begin{align*}
t \;\;::=\;\;
  &\; t_{\mathit{ret}}
  \;\mid\; t_{\mathbf{let}}(t')
  \;\mid\; t_{\mathbf{sample}}(v, t')
  \;\mid\; t_{\mathbf{commit}}(v, t')
  \;\mid\; t_{\mathbf{reveal}}(t')
\end{align*}
Deterministic steps ($\Let$, $\Reveal$) carry no choice data; $\Sample$ and
$\Commit$ record the value~$v$ chosen at that step.
\end{definition}

\begin{definition}[Trace Outcome and Weight]\label{def:traceweight}
Given a trace $t \in \Tr{p}$ and environment~$\rho$:
\begin{itemize}[nosep]
\item $\trout(t, \rho)$ follows the trace deterministically, evaluating
      expressions and extending the environment at each step, returning the
      final outcome.
\item $\trweight{\sigma}(t, \rho)$ computes the probability weight as a product:
\end{itemize}
\[
\trweight{\sigma}(t, \rho) =
\begin{cases}
  1 & \text{if } t = t_{\mathit{ret}} \\
  \trweight{\sigma}(t', \rho') & \text{if } t = t_{\mathbf{let}}(t')
    \text{ or } t = t_{\mathbf{reveal}}(t') \\
  D(\rho_m)(v) \cdot \trweight{\sigma}(t', \rho[x \mapsto v])
    & \text{if } t = t_{\mathbf{sample}}(v, t') \\
  \sigma(i,x,R)(\rho)(v)
    \cdot \trweight{\sigma}(t', \rho[x \mapsto v])
    & \text{if } t = t_{\mathbf{commit}}(v, t')
\end{cases}
\]
where $\rho'$ is the appropriately extended environment for the deterministic
step.
\end{definition}

\begin{definition}[Trace Legality]\label{def:tracelegal}
A trace is \emph{legal} if at every $\Sample$ step the value is in the
distribution's support, and at every $\Commit$ step the value satisfies the
guard.
\end{definition}

\begin{definition}[Reachability]\label{def:reach}
An outcome~$o$ is \emph{reachable} under profile~$\sigma$, written
$\Reach(\sigma, p, \rho, o)$, if there is a sequence of choices---each
sample value in the distribution's support, each commit value in the
kernel's support---leading from $(p, \rho)$ to outcome~$o$. This is
an inductive predicate mirroring the structure of $\outcomeDist$.

Separately, \emph{profile-free reachability} $\CanReach(p, \rho, o)$ requires
sample values in distribution support and commit values satisfying the guard
(legality), independent of any profile.
\end{definition}


\subsection{Key Theorems}\label{sec:theorems}

\begin{theorem}[Normalization]\label{thm:norm}
If $\NormDists(p)$ and $\NormOn(\sigma, p)$, then
\[
  \mathrm{totalWeight}(\densub{p}{\sigma}(\rho)) = 1.
\]
\leanref{outcomeDist\_totalWeight\_eq\_one}
\end{theorem}

\begin{theorem}[Adequacy]\label{thm:adequacy}
For every outcome~$o$:
\[
  \densub{p}{\sigma}(\rho)(o) \;=\;
  \sum_{t \in \Tr{p},\;\trout(t,\rho)=o} \trweight{\sigma}(t, \rho)
\]
\leanref{adequacy\_pointwise}
\end{theorem}

\begin{theorem}[Support Correctness]\label{thm:support}
An outcome~$o$ is in the support of $\densub{p}{\sigma}(\rho)$ if and only if
$\Reach(\sigma, p, \rho, o)$ holds
(\leanref{reach\_iff\_outcomeDist\_support}). Equivalently, $o$ is in the
support iff there exists a trace $t$ with $\trout(t,\rho) = o$ and
$\trweight{\sigma}(t,\rho) > 0$
(\leanref{mem\_support\_outcomeDist\_iff\_exists\_pos\_weight\_trace}).
\end{theorem}

\begin{theorem}[Fair Play]\label{thm:fairplay}
Under a fair-play operational profile~$\sigma$:
\begin{enumerate}[nosep]
\item Every trace with positive weight is legal
      (\leanref{admissible\_pos\_weight\_legal}).
\item Profile-dependent reachability implies profile-free reachability:
      $\Reach(\sigma, p, \rho, o) \Rightarrow \CanReach(p, \rho, o)$
      (\leanref{admissible\_reach\_canReach}).
\end{enumerate}
\end{theorem}

\noindent
Together with support correctness, this means that under fair play the
reachable outcomes of a Vegas program depend only on the program's guards and
distribution supports, not on the specific strategy profile.

\begin{theorem}[Commutativity]\label{thm:comm}
Adjacent commits with disjoint variable dependencies produce the same outcome
distribution regardless of their order. Adjacent reveals also commute:
\begin{align*}
  \densub{\Commit\;x\;\cdots\;\mathbf{in}\;\Commit\;y\;\cdots\;\mathbf{in}\;k}{\sigma}(\rho)
  &= \densub{\Commit\;y\;\cdots\;\mathbf{in}\;\Commit\;x\;\cdots\;\mathbf{in}\;k'}{\sigma}(\rho) \\
  \densub{\Reveal\;x'\;\cdots\;\mathbf{in}\;\Reveal\;y'\;\cdots\;\mathbf{in}\;k}{\sigma}(\rho)
  &= \densub{\Reveal\;y'\;\cdots\;\mathbf{in}\;\Reveal\;x'\;\cdots\;\mathbf{in}\;k'}{\sigma}(\rho)
\end{align*}
(under appropriate disjointness conditions on variables and
guard dependencies).

\noindent\leanref{outcomeDist\_comm\_commit},
\leanref{outcomeDist\_comm\_reveal}
\end{theorem}

\noindent
Commutativity confirms that the sequential textual order of independent
constructs is an artifact of the syntax, not a semantic distinction.


\subsection{Action Graph Semantics}\label{sec:actiongraph}

The denotational and trace semantics both follow the program's syntactic
structure. The \emph{action graph} semantics provides a complementary,
order-independent view: a Vegas program is flattened into a dependency DAG
of event nodes, and execution proceeds by firing any node whose dependencies
are satisfied.

\begin{definition}[Event Node and Action Graph]
An \emph{event node} has an identifier (the variable it binds), a kind
($\mathit{let}$, $\mathit{sample}$, $\mathit{commit}$, or $\mathit{reveal}$),
and a dependency list (the variable identifiers it reads). The function
$\mathit{extractEvents}(p)$ walks a Vegas program and emits one event node per
construct, with dependencies drawn from expression dependency tracking
($\mathsf{deps}$) for $\Let$ and $\Sample$ nodes, from the variable
identifiers in the player's view
$(\Gamma\!\!\downarrow_i).\mathit{vars}$ for $\Commit$ nodes, and from the
source variable for $\Reveal$ nodes.
\leanref{VegasCore.extractEvents}
\end{definition}

\begin{definition}[Configuration and Execution]
A \emph{configuration} consists of a list of pending event nodes and a partial
environment mapping variable identifiers to tagged values. A node is
\emph{ready} when all its dependencies are resolved. A \emph{step} fires a
ready node, recording a value and removing the node from the pending list. An
\emph{execution} is a sequence of steps from an initial configuration to a
terminal one (no pending nodes).
\leanref{Config}, \leanref{Step}, \leanref{Execution}
\end{definition}

The key property is that the final environment depends only on \emph{what}
values are assigned, not on the \emph{order} in which ready nodes are fired:

\begin{theorem}[Execution Uniqueness]\label{thm:execunique}
Two complete executions from the same initial configuration that assign the
same value at each node produce the same final environment.

\noindent\leanref{execution\_env\_unique}
\end{theorem}

\noindent
This result justifies the DAG perspective: the syntactic ordering of a Vegas
program is one valid topological sort of the action graph, but any other is
equally valid. The action graph also provides the conceptual foundation for
MAID compilation (\Cref{sec:maid}), where the DAG structure maps directly to
the MAID's graphical structure.


%% =====================================================================
\section{Strategic Bridge}\label{sec:strategic}
%% =====================================================================

The denotational semantics gives us a distribution over outcomes, but to apply
game-theoretic solution concepts we need a \emph{game structure}. The
strategic bridge provides this connection.

\subsection{Kernel Games}\label{sec:kernelgame}

\begin{definition}[Kernel Game]
A \emph{kernel game} $\KG$ over players~$\Players$ consists of:
\begin{itemize}[nosep]
\item $\KG.\mathrm{Strategy}(i)$: a strategy type for each player $i \in \Players$.
\item $\KG.\Outcome$: an outcome type.
\item $\KG.\mathrm{utility} : \KG.\Outcome \to (\Players \to \mathbb{R})$:
      a utility function.
\item $\KG.\mathrm{outcomeKernel} :
      (\forall i,\, \KG.\mathrm{Strategy}(i)) \to \PMF{\KG.\Outcome}$:
      a map from joint strategy profiles to outcome distributions.
\end{itemize}
Expected utility is
$\EU(\KG, \sigma, i) = \mathbb{E}_{o \sim \KG.\mathrm{outcomeKernel}(\sigma)}
  [\KG.\mathrm{utility}(o)(i)]$.
\end{definition}

\subsection{From Vegas to Kernel Games}\label{sec:tokernelgame}

Given $p : \Gamma \vdash p$, environment $\rho$, and proof $\NormDists(p)$,
we construct:
\[
  \toKG(p, \rho) : \KG \quad\text{where}\quad
  \begin{cases}
    \mathrm{Strategy}(i) &= \mathcal{B}_i \\
    \Outcome &= \Players \to_{\mathrm{fin}} \mathbb{Z} \\
    \mathrm{utility}(o)(i) &= o(i) \\
    \mathrm{outcomeKernel}(\sigma) &=
      (\densub{p}{\widehat{\sigma}}(\rho)).\mathrm{toPMF}
  \end{cases}
\]
Here $\mathcal{B}_i$ is the type of player~$i$'s view-based behavioral
strategies, and $\widehat{\sigma}$ is the induced operational profile
obtained by projecting each erased environment to the relevant player view.
The normalization theorem (\Cref{thm:norm}) ensures total weight~1, enabling
the promotion to $\PMF{}$.

\subsection{Expected Utility Preservation}

\begin{theorem}[EU Preservation]\label{thm:eu}
For any player~$i$ and joint strategy $\sigma$:
\[
  \EU(\toKG(p, \rho), \sigma, i) \;=\;
    \sum_{o \in \Outcome} \densub{p}{\widehat{\sigma}}(\rho)(o) \cdot o(i)
\]
\leanref{toKernelGame\_eu}
\end{theorem}

\subsection{Equilibrium Wrapper Layer}

Because the bridge preserves expected utility, game-theoretic solution concepts
apply directly to Vegas programs. We provide a thin wrapper layer that
transports the GT library's definitions through $\toKG$:

\begin{definition}[Vegas Nash Equilibrium]
A behavioral profile~$\sigma$ is a \emph{Nash equilibrium} of a Vegas
program~$p$ with environment~$\rho$ if it is a Nash equilibrium of
$\toKG(p, \rho)$:
\[
  \mathsf{IsNash}(p, \rho, \sigma) \;\;\iff\;\;
  \toKG(p, \rho).\mathsf{IsNash}(\sigma)
\]
\leanref{Vegas.IsNash}
\end{definition}

\noindent
The wrapper also provides $\mathsf{IsBestResponse}$, $\mathsf{IsDominant}$,
$\mathsf{IsParetoEfficient}$, and $\mathsf{socialWelfare}$, all defined by
the same transport. Two structural results lift through the bridge:

\begin{theorem}[Nash Characterization]\label{thm:nashchar}
$\mathsf{IsNash}(p, \rho, \sigma)$ holds iff every player is playing a best
response:
\[
  \mathsf{IsNash}(p, \rho, \sigma) \;\;\iff\;\;
  \forall\, i.\; \mathsf{IsBestResponse}(p, \rho, i, \sigma, \sigma_i)
\]
\leanref{Vegas.isNash\_iff\_bestResponse}
\end{theorem}

\begin{theorem}[Dominance implies Nash]\label{thm:domNash}
If every player plays a dominant strategy, the resulting profile is a Nash
equilibrium.

\noindent\leanref{Vegas.dominant\_is\_nash}
\end{theorem}

\noindent
For Matching Pennies, a Nash equilibrium is thus a pair of behavioral
strategies $(\sigma_{\pl{0}}, \sigma_{\pl{1}})$---each mapping the player's
visible state to a distribution over $\{\mathit{true}, \mathit{false}\}$---such
that neither player can increase their expected payoff by switching to a
different strategy. Since both commits precede all reveals, each player's
visible state at their commit point contains only their own prior
choices (if any), enforcing the simultaneity of the original game.


%% =====================================================================
\section{MAID Compilation}\label{sec:maid}
%% =====================================================================

Multi-Agent Influence Diagrams (MAIDs)~\cite{koller2003multi} are a
DAG-based representation of strategic interactions, extending Bayesian
networks with decision and utility nodes. We compile Vegas programs to a
restricted class of MAIDs and prove the compilation correct, including a
Kuhn theorem.

\subsection{Information Structure}\label{sec:maid-info}

Vegas's commit-reveal discipline gives rise to a particular information
structure: every value in the game is either \emph{fully public} (visible to
all players) or \emph{private to exactly one player}. There is no partial
sharing---no situation where player~$i$ observes a value but player~$j$ does
not, even though both have downstream decisions depending on it. This is
precisely the \emph{factored-observation} model of
Kova\v{r}\'ik et al.~\cite{kovarik2019rethinking}.

The connection to Vegas is direct:
\begin{itemize}[nosep]
\item $\Sample$ produces a chance value---immediately public.
\item $\Commit$ produces a hidden value---private to the committing player until
      explicitly revealed.
\item $\Reveal$ transitions a hidden value to public.
\item $\Let$ and $\Ret$ compute from already-visible values; they produce no new
      hidden state.
\end{itemize}
At any program point, a player's \emph{view} consists of all public values plus
their own hidden values. Because Vegas has no construct for partial disclosure
(revealing to some players but not others), the factored-observation property
is a structural consequence of the calculus.

\subsection{Compiled MAID Structure}\label{sec:maid-structure}

The compiled MAID inherits the factored-observation property. In a general
MAID, each decision node has two parent sets: data-dependency parents (values
the node's function reads) and observation parents (values the player actually
sees), with observation parents $\subseteq$ data parents. In a Vegas-compiled
MAID, these two sets are always \emph{equal}: a decision node depends only on
values visible to its player. The view-scoping judgment guarantees this by
construction.

\Cref{fig:compilation} shows the mapping from Vegas constructs to MAID nodes.
Only $\Sample$, $\Commit$, and $\Ret$ produce nodes; $\Let$ and $\Reveal$ are
compiled away through dependency tracking.

\begin{figure}[ht]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lll@{}}
\toprule
Vegas construct & MAID node & Notes \\
\midrule
$\Sample\;x \sim D$ & chance & CPD from $D$; parents from $\mathsf{deps}(D)$ \\
$\Commit\;x\;\mathbf{by}\;i\;\mathbf{where}\;R$
  & decision & Owner $i$; parents $= \mathsf{deps}(\Gamma\!\!\downarrow_i)$ \\
$\Ret$ & utility ($\times |\Players|$) & One per player; parents from payoff deps \\
$\Let\;x = e$ & \emph{(none)} & Deps inlined into downstream nodes \\
$\Reveal\;y\;\mathbf{from}\;x$ & \emph{(none)} & Updates visibility; no MAID node \\
\bottomrule
\end{tabular}
\caption{Vegas-to-MAID compilation.}\label{fig:compilation}
\end{figure}

\paragraph{Completeness (not mechanized).}
The compilation is also complete over factored-observation MAIDs, in the
following sense: every factored-observation MAID whose node functions are
expressible in~$L$ can be represented as a Vegas program that compiles to an
equivalent MAID.

The construction proceeds by topological sort. Process each node in order:
\begin{itemize}[nosep]
\item \emph{Chance node} $\to$ $\Sample$ with the node's CPD as the
      distribution expression.
\item \emph{Decision node} for player~$i$ $\to$ first emit $\Reveal$ for every
      unrevealed decision of a \emph{different} player that appears among this
      node's parents, then $\Commit$ with guard $\mathit{true}$.
\item \emph{Utility nodes} $\to$ $\Ret$ with the corresponding payoff functions.
\end{itemize}
The factored-observation invariant ensures the reveals are well-scheduled:
if another player's decision is a parent of this decision node, it must be
visible to the deciding player; since it belongs to a different player, it
must be public, meaning it was either a chance node (always public) or was
previously revealed. DAG acyclicity prevents circular dependencies. The
resulting program satisfies the standing assumptions: since MAIDs have no
action restrictions, the construction uses trivially-$\mathit{true}$ guards
(which satisfy $\Legal$, since every value passes the guard), and SSA
freshness follows from topological ordering.

Together with the compilation correctness theorem (\Cref{sec:maid-correct}),
this gives a precise characterization: Vegas programs capture exactly the
factored-observation MAIDs (up to expressiveness of~$L$). We have not
mechanized this direction; the formal development covers only the
compilation from Vegas to MAIDs.

\paragraph{Example: compiled MAIDs.}
For Matching Pennies, the compiler emits four nodes
(\Cref{fig:maid-mp}):

\begin{figure}[ht]
\centering
\fbox{\parbox{0.88\textwidth}{
\textbf{Matching Pennies} (simultaneous):
\[
\underset{\text{Player $\pl{0}$}}{\fbox{$D_a$}}
\qquad
\underset{\text{Player $\pl{1}$}}{\fbox{$D_b$}}
\qquad
\overset{D_a, D_b}{\longrightarrow}\;
\underset{\pl{0}}{\Diamond\; U_0}
\qquad
\overset{D_a, D_b}{\longrightarrow}\;
\underset{\pl{1}}{\Diamond\; U_1}
\]
Decision nodes $D_a$ (Player~$\pl{0}$) and $D_b$ (Player~$\pl{1}$) have
\emph{no observation parents}---neither player observes any prior node.
Both utility nodes $U_0, U_1$ have parents $\{D_a, D_b\}$.

\medskip
\textbf{Sequential Reveal} (Player~$\pl{0}$ moves first):
\[
\underset{\text{Player $\pl{0}$}}{\fbox{$D_a$}}
\;\xrightarrow{\;\text{obs}\;}\;
\underset{\text{Player $\pl{1}$}}{\fbox{$D_b$}}
\qquad
\overset{D_a, D_b}{\longrightarrow}\;
\underset{\pl{0}}{\Diamond\; U_0}
\qquad
\overset{D_a, D_b}{\longrightarrow}\;
\underset{\pl{1}}{\Diamond\; U_1}
\]
Now $D_b$ has observation parent $\{D_a\}$: the reveal of $a$ before
Player~$\pl{1}$'s commit makes $a$'s decision node observable. The
$\Reveal$ construct itself produces no MAID node; its effect is captured
entirely by the observation-parent edge.
}}
\caption{Compiled MAIDs for the two main examples. Boxes are decision
nodes, diamonds are utility nodes. Arrows labeled ``obs'' are observation
edges; other arrows are value-parent edges.}\label{fig:maid-mp}
\end{figure}


\subsection{Correctness}\label{sec:maid-correct}

The compilation correctness proof connects the MAID's evaluation to the
Vegas denotational semantics via two bridges:

\begin{enumerate}[nosep]
\item A \emph{reverse bridge} shows that for any MAID policy, evaluating the
      compiled MAID and extracting the outcome distribution equals the Vegas
      behavioral outcome distribution under the reflected policy. The
      reflected policy translates a MAID policy---which maps infosets to action
      distributions---back into a Vegas behavioral profile by reconstructing
      each player's view from the MAID's parent configuration.

\item A \emph{pure bridge} shows that for any Vegas pure strategy profile,
      compiling it to a MAID pure policy and reflecting back recovers the
      original pure outcome distribution. The key step is showing that the
      view reconstruction is faithful: the player's projected view under the
      MAID's raw-node environment agrees with the view under the Vegas
      typed environment.
\end{enumerate}

\noindent
The pure bridge requires a \emph{view-insensitivity} argument: each variable's
contribution to the environment depends only on the MAID nodes in its
dependency set, so extending the raw environment at unrelated positions does
not alter the player's view. This is the most technically demanding part of
the proof, involving a witness construction that maps the Vegas environment
back to a MAID type-assignment and shows agreement at all view-dependency
positions.

\subsection{Perfect Recall}\label{sec:perfect-recall}

The compiled MAID has \emph{perfect recall}: each decision node's observation
parents include all nodes previously observed by the same player. This follows
from the compilation strategy together with the reveal-time tracking described
above. Because Vegas programs execute top-to-bottom and the compiler records
observation parents as $\Gamma\!\!\downarrow_i$ (the player's full view at the
point of commitment), any previously revealed variable is already in the view.
For variables hidden from the player, the reveal-time invariant ensures that
the corresponding node was not yet public at the time of the earlier decision.

\subsection{The Vegas Kuhn Theorem}\label{sec:kuhn}

Combining the compilation correctness bridges with perfect recall yields a
Kuhn-style result for Vegas programs.

\begin{theorem}[Vegas Kuhn Theorem]\label{thm:kuhn}
Let $p$ be a Vegas program satisfying the standing assumptions (\Cref{def:standing}).
For every product mixture $\mu$ over pure strategy profiles (one independent
distribution per player over that player's pure strategies), there exists a
behavioral strategy profile $\sigma$ such that:
\[
  \outcomeDist_{\sigma}(p, \rho) \;=\;
  \mathbb{E}_{\pi \sim \mu}\bigl[\outcomeDist_{\pi}(p, \rho)\bigr]
\]
\end{theorem}

\noindent
The proof compiles $p$ to a MAID, applies the MAID-level Kuhn theorem
(which requires perfect recall) to obtain a MAID behavioral policy, and
reflects this policy back to a Vegas behavioral profile via the reverse bridge.
The pure bridge establishes that the right-hand side---the mixture over
pure outcome distributions---equals the corresponding MAID-level mixture,
closing the chain.

This theorem is significant because it shows that behavioral strategies
suffice: any outcome distribution achievable by mixing over pure strategies
can be realized by a single behavioral profile. In game-theoretic terms, there
is no loss of generality in restricting attention to behavioral strategies
when analyzing Vegas programs.


%% =====================================================================
\section{Mechanization}\label{sec:mechanization}
%% =====================================================================

The entire development is formalized in Lean~4, building on Mathlib for
mathematical foundations (measure theory, probability mass functions,
finite-support functions). \Cref{tab:loc} summarizes the codebase.

\begin{table}[ht]
\centering
\begin{tabular}{lrr}
\toprule
Component & Files & Lines \\
\midrule
Vegas core (types, contexts, constructs) & 2 & ${\sim}650$ \\
Vegas well-formedness, scoping \& assumptions & 2 & ${\sim}580$ \\
Vegas strategies \& operational profiles & 2 & ${\sim}290$ \\
Vegas semantics (denotational, trace) & 3 & ${\sim}870$ \\
Vegas strategic bridge \& equilibrium & 2 & ${\sim}150$ \\
Vegas MAID compilation, bridges \& Kuhn & 8 & ${\sim}6{,}300$ \\
Vegas examples & 1 & ${\sim}250$ \\
FDist library & 1 & ${\sim}400$ \\
Supporting expression language & 2 & ${\sim}580$ \\
Action graph & 1 & ${\sim}390$ \\
\midrule
\textbf{Vegas subtotal} & \textbf{25} & $\mathbf{{\sim}10{,}500}$ \\
\midrule
GameTheory library & 60+ & ${\sim}33{,}000$ \\
\midrule
\textbf{Total} & & $\mathbf{{\sim}43{,}000}$ \\
\bottomrule
\end{tabular}
\caption{Lines of Lean~4 code by component (approximate).}\label{tab:loc}
\end{table}

\paragraph{Expression language parametricity.}
The Vegas calculus is generic over its expression language interface
$L : \leanref{IExpr}$. The interface bundles:
\begin{itemize}[nosep]
\item A type universe $L.\Ty$ with distinguished $L.\bool$ and $L.\intt$.
\item A value family $L.\Val(\tau)$ and expression syntax
      $L.\Expr(\Delta, \tau)$ with an evaluator.
\item Distribution expressions $L.\mathit{DistExpr}(\Delta, \tau)$ evaluating
      to weighted distributions $\FDist{L.\Val(\tau)}$.
\item Dependency tracking: $L.\mathit{exprDeps}$ and $L.\mathit{distDeps}$
      assign each expression the set of variable identifiers it depends on.
\item Soundness laws: evaluation depends only on tracked variables. These are
      required for the commutativity proofs (\Cref{thm:comm}), which must show
      that reordering independent constructs does not affect expression
      evaluation.
\item Structural laws: expressions can be extended across fresh bindings, and
      expressions not depending on a variable can be re-expressed after
      dropping it. These are required for guard manipulation in the
      commutativity proofs.
\end{itemize}
The metatheory is proved once for any $L$. A concrete instantiation with
integers, Booleans, arithmetic, equality, and conditionals
(\leanref{simpleExpr}) is used for examples and testing.

\paragraph{Design decisions.}
\begin{itemize}[nosep]
\item \emph{FDist vs PMF.} We use $\FDist{\alpha} = \alpha \to_{\mathrm{fin}} \NNQ$
      (finitely-supported nonneg-rational functions) as the primary distribution
      type, rather than Mathlib's $\PMF{}$. This avoids carrying normalization
      proofs through intermediate computations; normalization is established once
      at the end (\Cref{thm:norm}) and used to promote to $\PMF{}$ when needed.

\item \emph{SSA via Fresh.} Well-formedness is a simple ``fresh identifier in
      context'' check rather than a full de~Bruijn encoding. This keeps the
      syntax readable while still providing the uniqueness properties needed for
      the MAID compilation (SSA structure maps directly to a DAG).
\end{itemize}


%% =====================================================================
\section{Related Work}\label{sec:related}
%% =====================================================================

\paragraph{Game description languages.}
GDL (Game Description Language) and its extensions provide logic-based
languages for describing games, primarily for general game playing. Gambit and
OpenSpiel offer computational frameworks for game solving. These focus on
execution and solving rather than verified semantics or type-based information
control.

\paragraph{Mechanized game theory.}
Prior work has mechanized game-theoretic results in Coq and Isabelle, including
Nash's theorem and various solution concepts. Our GameTheory library in Lean~4
provides extensive-form games, normal-form games, and MAIDs with conversions
between them, but the primary contribution here is the Vegas calculus and its
verified semantics, not the GT library itself.

\paragraph{Information flow and security types.}
Vegas's visibility annotations are reminiscent of information-flow type systems
in the security literature. Both track which principals can observe which data.
The key difference is purpose: security type systems prevent information
\emph{leakage}, while Vegas's view-scoping discipline enforces game-theoretic information
\emph{structure}---ensuring that strategies depend only on the information
players are supposed to have.

\paragraph{Blockchain and commit-reveal.}
The commit-reveal pattern is standard in blockchain applications (e.g.,
sealed-bid auctions on Ethereum). Vegas abstracts this pattern into a
view-scoping discipline, showing that the same information-structure guarantees
can be stated at the language level rather than deferred entirely to cryptography. This complements work
on verified smart contracts by providing a game-theoretic foundation for
reasoning about strategic interaction in decentralized protocols.

\paragraph{Probabilistic programming.}
Vegas shares structural similarities with probabilistic programming languages
(sampling, monadic binding), but its focus is fundamentally different:
probabilistic programs model a single agent's uncertainty, while Vegas models
multi-agent strategic interaction with per-player information constraints.


%% =====================================================================
\section{Conclusion}\label{sec:conclusion}
%% =====================================================================

We have presented Vegas, a typed game calculus that integrates imperfect
information structure through commit-reveal constructs, visibility-aware
contexts, and a view-scoping discipline. The calculus comes with four
semantic representations---denotational, trace, action graph, and
native---connected by adequacy, normalization, execution uniqueness, and
commutativity theorems. A strategic bridge to kernel games preserves expected
utility, enabling game-theoretic solution concepts; a verified compilation
path to MAIDs preserves outcome distributions. The entire development is
mechanized in Lean~4.

\paragraph{Limitations.}
The current calculus is restricted to finite types (all $L.\Val(\tau)$ are
finite), which excludes games with continuous action spaces. The commit-reveal
model assumes perfect recall---a player's view grows monotonically as the
program executes, and there is no construct for forgetting. The expression
language interface requires explicit dependency tracking with soundness
proofs, which adds friction to defining new expression languages.

\paragraph{Future directions.}
Extending Vegas with richer expression languages (continuous distributions,
recursive types), sequential equilibrium refinements, and a deeper connection
to verified smart contract languages. The commit-reveal structure suggests a
natural correspondence with zero-knowledge protocols that remains to be
formalized.

\begin{thebibliography}{9}

\bibitem{koller2003multi}
D.~Koller and B.~Milch.
\newblock Multi-agent influence diagrams for representing and solving games.
\newblock {\em Games and Economic Behavior}, 45(1):181--221, 2003.

\bibitem{kovarik2019rethinking}
V.~Kova\v{r}\'ik, M.~Schmid, N.~Burch, M.~Bowling, and V.~Lis\'y.
\newblock Rethinking formal models of partially observable multiagent decision
  making.
\newblock {\em Artificial Intelligence}, 277:103171, 2019.

\end{thebibliography}

\end{document}
